Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-07-02 manuscript/pnas_style/manuscript.tex
moved sentence

Our analyses reveal that a sizable proportion of allele frequency change in
\vb{these experimental evolution populations} is due to the (likely indirect)
action of selection. Capitalizing on replicated designs, we characterized the
extent to which convergent selection pressures lead to parallel changes in
allele frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations. Finally, \vb{through
  simulation results, we show that for a given level of additive genetic
  variance, the strengths of temporal and replicate covariances depend on the
  mode of selection, the details of the populations or selection experiment,
  and the level of linkage disequilibria, yet the level of temporal covariance
  is relatively invariant to the number of loci underlying fitness, as long as
fitness is sufficiently polygenic.}

These estimates of the contribution of selection could be refined by using
patterns of linkage disequilibria (LD) and recombination which would allow
us to more fully parameterize a linked-selection model of temporal allele
frequency change \cite{Buffalo2019-io}. The basic prediction is that
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. However, one limitation of these pooled sequence datasets is
that none of the studies we reanalyzed estimated linkage disequilibria data for
the evolved populations.  While there are LD data for a natural population of
\emph{D. simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a
relationship between temporal covariance and LD.  We believe this is driven by
the idiosyncratic nature of LD in evolve-and-resequence populations, which
often extends over large genomic distances
\cite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
and recombination maps would allow one to disentangle the influence of closely
linked sites from more distant sites in causing temporal autocovariance, and
allow the fitting of more parametric models to estimate population parameters
such as the additive genetic variance for fitness directly from temporal
genomic data alone \cite{Buffalo2019-io}.

>>>>

Our analyses reveal that a sizable proportion of allele frequency change in
\vb{these experimental evolution populations} is due to the (likely indirect)
action of selection. Capitalizing on replicated designs, we characterized the
extent to which convergent selection pressures lead to parallel changes in
allele frequencies across replicate populations, and found that a substantial
proportion of the response is shared across short timescales.  These likely
represent substantial under-estimates of the contribution of linked selection
because the studies we have reanalyzed do not sequence the population each
generation, preventing us from including the effects of stronger correlations
between adjacent generations. Furthermore, our estimation methods are
intentionally conservative, for example they exclude the contribution of
selection that does not persist across generations and selection that reverses
sign; thus they can be seen as a lower bound of the effects of selection, which
we have confirmed through forward-in-time simulations.  \vb{Finally, through
  simulation results, we show that for a given level of additive genetic
  variance, the strengths of temporal and replicate covariances depend on the
  mode of selection, the details of the populations or selection experiment,
  and the level of linkage disequilibria, yet the level of temporal covariance
  is relatively invariant to the number of loci underlying fitness, as long as
fitness is sufficiently polygenic.}

These estimates of the contribution of selection could be refined by using
patterns of linkage disequilibria (LD) and recombination which would allow us
to more fully parameterize a linked-selection model of temporal allele
frequency change \cite{Buffalo2019-io}. The basic prediction is that regions of
higher linkage disequilibrium and lower recombination should have greater
temporal autocovariance than regions with lower LD and higher recombination.
However, one limitation of these pooled sequence datasets is that none of the
studies we reanalyzed estimated linkage disequilibria data for the evolved
populations.  While there are LD data for a natural population of \emph{D.
simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a relationship
between temporal covariance and LD.  We believe this is driven by the
idiosyncratic nature of LD in evolve-and-resequence populations, which often
extends over large genomic distances \cite{Nuzhdin2013-gf,Kelly2019-dc}. Future
studies complete with LD data and recombination maps would allow one to
disentangle the influence of closely linked sites from more distant sites in
causing temporal autocovariance, and allow the fitting of more parametric
models to estimate population parameters such as the additive genetic variance
for fitness directly from temporal genomic data alone \cite{Buffalo2019-io}.
\vb{Future work could refine our $G(t)$ estimates by including selection's
  impact on the variance in allele frequency terms (e.g. see equation 26 of
\cite{Buffalo2019-io}), and possibly quantifying the covariances missed when
sequencing is not done each generation; both would lead to less conservative
estimates that could show a large impact of selection. }
